Irreducible Components and Dimension of the Springer Fiber of a Hook-Type Slodowy Slice

### 2024-09-03 paper.tex
up to 9.2

\subsection{The Centralizer of a Nilpotent Matrix}
\begin{definition}
    A matrix \(Y\) is \emph{Toeplitz} if it is constant along bands parallel to the main diagonal.
    That is, \(\forall i,j,k. \; Y_{ij} = Y_{i + k, j + k}\).
\end{definition}
\begin{definition}
    An \(m \times n\) matrix \(Y\) is \emph{lower-left Toeplitz} if it is Toeplitz and, in addition, we have \(y_{n - i, j - 1} = 0\) whenever \(i + j \geq \min(m,n)\).
\end{definition}
That is, \(Y\) is lower-left Toeplitz if it is Toeplitz, and the only nonzero entries are those with Manhattan distance less than \(\min(m,n)\) from the entry in the bottom-left corner.
In yet other words, all but the leftmost (equivalently, bottommost) \(\min(m,n)\) diagonal bands are zero.

\begin{lemma}\label{nilpotent_centralizer}
    Let \(\lambda = (\lambda_1, ..., \lambda_k)\) be a partition of \(m\).
    The centralizer of \(J(\lambda)\) in \(\gl_m\) is the subalgebra consisting of matrices 
    \[M = \begin{pmatrix}
        M_{11} & \cdots & M_{1k}\\
        \vdots &        & \vdots\\
        M_{k1} & \cdots & M_{kk}
    \end{pmatrix},\]
    where each \(M_{ij}\) is a \(\lambda_i \times \lambda_j\) matrix, such that each \(M_{ij}\) is lower-left Toeplitz.
\end{lemma}
\begin{proof}
    Let 
    \[M = \begin{pmatrix}
        M_{11} & \cdots & M_{1k}\\
        \vdots &        & \vdots\\
        M_{k1} & \cdots & M_{kk}
    \end{pmatrix}.\]
    We need to show that \(J(\lambda)M = MJ(\lambda)\) if and only if each \(M_{ij}\) is lower-left Toeplitz.
    \par We have 
    \[J(\lambda)M = \begin{pmatrix}
        J_{\lambda_1}M_{11} & \cdots & J_{\lambda_1}M_{1k}\\
        \vdots & & \vdots\\
        J_{\lambda_k}M_{k1} & \cdots & J_{\lambda_k}M_{kk}
    \end{pmatrix}, \textrm{ and } MJ(\lambda) = \begin{pmatrix}
        M_{11} J_{\lambda_1} & \cdots & M_{1k} J_{\lambda_k}\\
        \vdots & & \vdots\\
        M_{k1} J_{\lambda_1} & \cdots & M_{kk} J_{\lambda_k}
    \end{pmatrix}.\]
    So, we have \(J(\lambda)M = MJ(\lambda)\) if and only if \(\forall i,j. \; J_{\lambda_i} M_{ij} = M_{ij} J_{\lambda_j}\).
    Multiplying on the left by \(J_{\lambda_i}\) just shifts each row down by one, and multiplying on the right by \(J_{\lambda_j}\) shifts each column left by one.
    The matrices for which left-shifting gives the same result as down-shifting are exactly the lower-left Toeplitz matrices.
\end{proof}

>>>>

\subsection{The Centralizer of a Nilpotent Matrix}
\begin{definition}
    A matrix \(Y\) is \emph{Toeplitz} if it is constant along bands parallel to the main diagonal.
    That is, \(Y\) is Toeplitz if for all \(i,j,k\) we have \(Y_{ij} = Y_{i + k, j + k}\).
\end{definition}
\begin{definition}
    An \(m \times n\) matrix \(Y\) is \emph{lower-left Toeplitz} if it is Toeplitz and, in addition, for all \(i\) and \(j\) with \(i + j \geq \min(m,n)\), we have \(y_{n - i, j - 1} = 0\).
\end{definition}
That is, \(Y\) is lower-left Toeplitz if it is Toeplitz, and the only nonzero entries are those with Manhattan distance less than \(\min(m,n)\) from the entry in the bottom-left corner.
In yet other words, all but the leftmost (equivalently, bottommost) \(\min(m,n)\) diagonal bands are zero.

\begin{lemma}\label{nilpotent_centralizer}
    Let \(\lambda = (\lambda_1, ..., \lambda_\ell)\) be a partition of \(m\).
    The centralizer of \(J(\lambda)\) in \(\gl_m\) is the subalgebra consisting of the matrices 
    \[M = \begin{pmatrix}
        M_{11} & \cdots & M_{1\ell}\\
        \vdots &        & \vdots\\
        M_{\ell 1} & \cdots & M_{\ell\ell}
    \end{pmatrix},\]
    where each \(M_{ij}\) is a \(\lambda_i \times \lambda_j\) matrix, such that each \(M_{ij}\) is lower-left Toeplitz.
\end{lemma}
\begin{proof}
    Let 
    \[M = \begin{pmatrix}
        M_{11} & \cdots & M_{1\ell}\\
        \vdots &        & \vdots\\
        M_{\ell 1} & \cdots & M_{\ell\ell}
    \end{pmatrix}.\]
    We need to show that \(J(\lambda)M = MJ(\lambda)\) if and only if each \(M_{ij}\) is lower-left Toeplitz.
    \par We have 
    \begin{align*}
        J(\lambda)M = \begin{pmatrix}
            J_{\lambda_1}M_{11} & \cdots & J_{\lambda_1}M_{1\ell}\\
            \vdots & & \vdots\\
            J_{\lambda_\ell}M_{\ell 1} & \cdots & J_{\lambda_\ell}M_{\ell\ell}
        \end{pmatrix} & , \textrm{ and} \\
        MJ(\lambda) = \begin{pmatrix}
            M_{11} J_{\lambda_1} & \cdots & M_{1\ell} J_{\lambda_\ell}\\
            \vdots & & \vdots\\
            M_{\ell 1} J_{\lambda_1} & \cdots & M_{\ell\ell} J_{\lambda_\ell}
        \end{pmatrix} & .
    \end{align*}
    So, \(J(\lambda)M = MJ(\lambda)\) if and only if for all \(i\) and \(j\) we have \(J_{\lambda_i} M_{ij} = M_{ij} J_{\lambda_j}\).
    Multiplying on the left by \(J_{\lambda_i}\) just shifts each row down by one, and multiplying on the right by \(J_{\lambda_j}\) shifts each column left by one.
    The matrices for which left-shifting gives the same result as down-shifting are exactly the lower-left Toeplitz matrices.
\end{proof}
